Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-16 prop_letter.tex
prop_letter.tex Dealing with the table height again...
prop_letter.tex \section{Gleason} Updating...

While the finite-precision extension of Kochen-Specker theorem asserts
the non-existence of certain kind of QIVPMs, we also want to understand
what a QIVPM could be when it exists. In Sec.~\ref{sec:Interval-Uncertainty},
the only non-trivial QIVPM~ $\bar{\mu}_{3}$ is induced by the quantum
probability measure~$\mu_{1}$ defined in Sec.~\ref{sec:fuzzy}.
This result will first be generalized in this section, and then we
will show some QIVPMs may not be induced by any quantum probability
measure, i.e., no state is consistent with these kind of QIVPMs.

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. Replacing numbers by intervals containing
them can be generalized if we carefully pick a set of intervals and
the rule to select the correct interval when intervals are overlapping.
For example, consider the following collection of intervals: 
\begin{eqnarray}
\mathscr{I}_{n} & = & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
with length~$\left(2n\right)^{-1}$ at each end. Since these intervals~$[\ell,r]$
are overlapping, we define the non-overlapping interior part of a
interval, $\textrm{Int}\left([\ell,r]\right)$, as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}
\begin{tabular}{cccc}
\toprule 
\multicolumn{2}{c}{$\mathscr{I}_{n}$} & \multicolumn{2}{c}{$\mathscr{I}_{2}$}\tabularnewline
\midrule
\midrule 
$[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$ & $[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline
\midrule 
$\imposs$ & $\imposs$ & $\imposs$ & $\imposs$\tabularnewline
\midrule 
$\left[0,\tfrac{1}{2n}\right]$ & $\left(0,\tfrac{0.5}{2n}\right]$ & $\left[0,\tfrac{1}{4}\right]$ & $\left(0,\tfrac{1}{8}\right]$\tabularnewline
\midrule 
$\begin{gathered}\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\\
\textrm{for odd }k
\end{gathered}
$ & $\left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$\tabularnewline
\midrule 
\multirow{2}{*}{$\begin{gathered}\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\\
\textrm{for even }k\\
0<k<2n
\end{gathered}
$} & \multirow{2}{*}{$\left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]$} & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$\tabularnewline\addlinespace
\cmidrule{3-4} 
 &  & $\left[\tfrac{1}{2},1\right]$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$\tabularnewline\addlinespace
\midrule 
$\left[\tfrac{2n-1}{2n},1\right]$ & $\left[\tfrac{2n-0.5}{2n},1\right)$ & $\left[\tfrac{1}{4},1\right]$ & $\left[\tfrac{7}{8},1\right)$\tabularnewline
\midrule 
$\necess$ & $\necess$ & $\necess$ & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
Hence, given a real-valued probability~$x$, we could easily replace
it with an interval~$\iota_{n}\left(x\right)=[\ell,r]$ if $x\in\textrm{Int}\left([\ell,r]\right)$.

After we have a rule to replace a real number by an interval, any
Born rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space can be extended to a QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
where $\circ$ is the function concatenation. For example, when $n=2$,
two quantum probability measure~$\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ can induce two QIVPMs, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for projector $P$ in the minimal subspace of events containing $\proj{0}$,
$\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed straightforwardly
in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}
\begin{tabular}{ccccc}
\toprule 
$P$ & $\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$ & $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\imposs$ & $0$ & $\imposs$\tabularnewline
\midrule 
$\proj{0}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{7}{24}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{3}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{17}{24}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\necess$ & $1$ & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
Verifying $\iota_{2}\circ\mu_{1}$ and $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
are QIVPMs are also straightforward.

A state~$\rho$ is called totally mixed state if it satisfies $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.
Given any totally mixed state~$\rho$, we know not only $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, but also the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need to define a notion
of distance based on the following norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $1/n$. Reversely, if the distance between $\rho$ and
$\rho'$ is bounded by $\left(12n\right)^{-1}$, then $\rho'$ is
consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$. For example,
$\mathbb{1}/3$ and $\rho_{0}$ are both totally mixed states. Since
$\left\Vert \frac{\mathbb{1}}{3}-\rho_{0}\right\Vert \le\frac{1}{24}$,
we should have $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}},\events\right)$
and $\mathbb{1}/3\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the states very close to
$\rho$ is guarantied belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.

>>>>

While the finite-precision extension of Kochen-Specker theorem asserts
the non-existence of certain kind of QIVPMs, we also want to understand
what a QIVPM could be when it exists. In Sec.~\ref{sec:Interval-Uncertainty},
the only non-trivial QIVPM~ $\bar{\mu}_{3}$ is induced by the quantum
probability measure~$\mu_{1}$ defined in Sec.~\ref{sec:fuzzy}.
This result will first be generalized in this section, and then we
will show some QIVPMs may not be induced by any quantum probability
measure, i.e., no state is consistent with these kind of QIVPMs, while
a quantum probability measure is always induced by an unique state
by Gleason's theorem.

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. Replacing numbers by intervals containing
them can be generalized if we carefully pick a set of intervals and
a rule to select the correct interval when intervals are overlapping.
For example, consider the following collection of intervals: 
\begin{eqnarray}
\mathscr{I}_{n} & = & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
with length~$\left(2n\right)^{-1}$ at each end. Since these intervals~$[\ell,r]$
are overlapping, we define the non-overlapping interior part of a
interval, $\textrm{Int}\left([\ell,r]\right)$, as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}

\begin{tabular}{cccc}
\toprule 
\addlinespace
\multicolumn{2}{c}{$\mathscr{I}_{n}$} & \multicolumn{2}{c}{$\mathscr{I}_{2}$}\tabularnewline\addlinespace
\midrule 
\addlinespace
$[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$ & $[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\imposs$ & $\imposs$ & $\imposs$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\left[0,\tfrac{1}{2n}\right]$ & $\left(0,\tfrac{0.5}{2n}\right]$ & $\left[0,\tfrac{1}{4}\right]$ & $\left(0,\tfrac{1}{8}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\begin{gathered}\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\\
\textrm{for odd }k
\end{gathered}
$ & $\left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
\multirow{2}{*}{$\begin{gathered}\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\\
\textrm{for even }k\\
0<k<2n
\end{gathered}
$} & \multirow{2}{*}{$\left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]$} & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$\tabularnewline\addlinespace
\cmidrule{3-4} 
\addlinespace
 &  & $\left[\tfrac{1}{2},1\right]$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\left[\tfrac{2n-1}{2n},1\right]$ & $\left[\tfrac{2n-0.5}{2n},1\right)$ & $\left[\tfrac{1}{4},1\right]$ & $\left[\tfrac{7}{8},1\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\necess$ & $\necess$ & $\necess$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Hence, given a real-valued probability~$x$, we could easily replace
it with an interval~$\iota_{n}\left(x\right)=[\ell,r]$ if $x\in\textrm{Int}\left([\ell,r]\right)$.

After we have the rule to replace a real number by an interval, any
Born rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space can be extended to a QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
where $\circ$ is the function concatenation~\citep{HOSTunpublished}.
For example, when $n=2$, two quantum probability measures~$\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ can induce two QIVPMs~$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for any projector $P$ in the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed
straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}

\begin{tabular}{ccccc}
\toprule 
\addlinespace
$P$ & $\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$ & $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$ & $0$ & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{0}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{7}{24}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{1}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{2}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{3}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{2}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{1}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{0}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{17}{24}$ & $\left[\frac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$ & $1$ & $\necess$ & $1$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Verifying $\iota_{2}\circ\mu_{1}$ and $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
are QIVPMs are also straightforward.

A state~$\rho$ is called totally mixed state if it satisfies $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.
Given any totally mixed state~$\rho$, we know not only $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, but also the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need to define a notion
of distance based on the following norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $1/n$. Reversely, if the distance between $\rho$ and
$\rho'$ is bounded by $\left(12n\right)^{-1}$, then $\rho'$ is
consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$~\citep{HOSTunpublished}.
For example, $\mathbb{1}/3$ and $\rho_{0}$ are both totally mixed
states. Since $\left\Vert \frac{\mathbb{1}}{3}-\rho_{0}\right\Vert \le\frac{1}{24}$,
we should have $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}},\events\right)$
and $\mathbb{1}/3\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the states very close to
$\rho$ is guarantied belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.
